---
title: "MuM_PreDiCT WP4 Polypharmacy (3) Pregnancy Induced Hypertension - Outcome Identification"
output: html_notebook
---

last updated: 2023.05.17 (Lisa Kent)

This notebook:
* Takes a three stage approach to identifying PIH

* 1. From medications (EPD)
* 2. From diagnosis code (PAS)
* 3. From "PRESENT PREGNANCY PROBLEMS" variable (NIMATS)

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```

__Lode odbc (Open Database Connectivity) package__
```{r}
if(!require(odbc)) install.packages("odbc")
library(odbc)
```

__Declare connection to database__
```{r}
con <- dbConnect(odbc::odbc(), "hirusql3", timeout = 10)
```

```{r}
#read in previous preg level data (if needed)
#mumpredict_wp4 <- read_csv
```




__OUTCOME: PIH__
__(1) EPD -> PIH defined using medications of interest__
```{r}
wp4_epd_outcome <- left_join(wp4_epd %>% select(STUDY_ID, VTM_NM, ISSUE_DATE), wp4_dates %>% select(STUDY_ID, PREG_ID, DATE_20W_GEST, DATE_6MO_POSTPART), by = "STUDY_ID", relationship = "many-to-many")
```


```{r}
#flag pregnancies with PIH-related medications

wp4_epd_outcome <- wp4_epd_outcome %>%
  mutate(
    PIH_OUTCOME_MED = if_else(
      grepl("amlodipine|doxazosin|labetalol|methyldopa|nifedipine", tolower(VTM_NM)) & 
        ISSUE_DATE >= DATE_20W_GEST & 
        ISSUE_DATE < DATE_6MO_POSTPART, 
      1, 0
    )
  )
```


```{r}
wp4_epd_outcome <- wp4_epd_outcome %>% select(PREG_ID, PIH_OUTCOME_MED) %>% distinct() %>% group_by(PREG_ID) %>% slice_max(PIH_OUTCOME_MED)
```


__(2) PAS -> PIH defined using ICD-10 codes of interest__
```{r}
pas <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[PAS_1]")
colnames(pas) <- toupper(colnames(pas))
```

```{r}
pas <- pas %>% 
  mutate(across(where(is.character), str_trim))
```


```{r}
pas_icd <- pas %>% select(STUDY_ID, ADMISSION_DATE, contains("DIAGNOSIS"))
```

```{r}
#create long df of ICD10 codes (with study-id and admission date)
pas_icd_long <- pas_icd %>%
  pivot_longer(cols = PRIMARY_DIAGNOSIS_CODE:SECONDARY_DIAGNOSIS_15, names_to = "VAR", values_to = "ICD10_CODE")
head(pas_icd_long, 20)
nrow(pas_icd_long)
```

```{r}
#remove rows where no ICD10 code recorded
pas_icd_long_narm <- pas_icd_long[!is.na(pas_icd_long$ICD10_CODE), ]
head(pas_icd_long_narm, 20)
nrow(pas_icd_long_narm)
```



```{r}
icd_outcomes <- read_csv("S:/HBS_E062/2. CODE/code lists/ICD_outcomes_full_2023.06.28.csv")
```

```{r}
#extract icd-10 codes related to pre-eclampsia and gestational hypertension
icd_pih_list <- icd_outcomes %>% filter(
  OUTCOME == "Preeclampsia_Prevalence_mumpredict_ICD10" |
    OUTCOME == "GestationalHypertension_mumpredict_ICD10") %>% 
  pull(READ_CODE)
```

```{r}
# append hypertension ICD10 codes to list
hypertension <- c("I11","I12","I13","I15","I10")
icd_pih_list <- append(icd_pih_list, hypertension, length(icd_pih_list))
```



```{r}
# for 4 characters ICD-10 code we extract the first 5 chars (i.e. XXX.X)
pas_icd_long_narm$ICD10_CODE_4CHAR <- substr(pas_icd_long_narm$ICD10_CODE, 1, 5)

# for 3 characters ICD-10 code we extract the first 3 chars (i.e. XXX)
pas_icd_long_narm$ICD10_CODE_3CHAR <- substr(pas_icd_long_narm$ICD10_CODE, 1, 3)
```

```{r}
wp4_pas_outcome <- left_join(pas_icd_long_narm %>% select(STUDY_ID, ADMISSION_DATE, ICD10_CODE_4CHAR, ICD10_CODE_3CHAR), wp4_dates %>% select(STUDY_ID, PREG_ID, DATE_20W_GEST, DATE_6MO_POSTPART), by = "STUDY_ID", relationship = "many-to-many")
```


```{r}
wp4_pas_outcome <- wp4_pas_outcome %>% mutate(
  PIH_OUTCOME_ICD = if_else(
    (ICD10_CODE_4CHAR %in% icd_pih_list | ICD10_CODE_3CHAR %in% icd_pih_list) &
      ADMISSION_DATE >= DATE_20W_GEST & 
        ADMISSION_DATE < DATE_6MO_POSTPART,
    1, 0
  )
)

```

```{r}
wp4_pas_outcome <- wp4_pas_outcome %>% select(PREG_ID, PIH_OUTCOME_ICD) %>% distinct() %>% group_by(PREG_ID) %>% slice_max(PIH_OUTCOME_ICD)
```



__(3) NIMATS -> PIH defined using variable "PRESENT PREGNANCY PROBLEMS"__
```{r}
wp4_pih_outcome <- wp4 %>% select(STUDY_ID, PREG_ID, `PRES_PREG_PROB_P.I.H.`, `PRES_PREG_PROB_P.I.H._+_PROTEINURIA`, PRES_PREG_PROB_ECLAMPSIA)
names(wp4_pih)
```

```{r}
wp4_pih_outcome <- wp4_pih_outcome %>%
  mutate(
    PIH_OUTCOME_NIMATS = if_else(`PRES_PREG_PROB_P.I.H.` == 1 |
                                   `PRES_PREG_PROB_P.I.H._+_PROTEINURIA` == 1 |
                                   `PRES_PREG_PROB_ECLAMPSIA` == 1,
                                 1, 0
                                 )
    )
```


__JOIN OUTCOME DATAFRAMES TOGETHER__
```{r}
wp4_pih_outcome <- left_join(wp4_pih_outcome, wp4_epd_outcome, by = "PREG_ID")
```

```{r}
wp4_pih_outcome <- left_join(wp4_pih_outcome, wp4_pas_outcome, by = "PREG_ID")
```


__CREATE OUTCOME FLAG WHICH COMBINES ALL THREE METHODS OF IDENTIFYING PIH__
```{r}
wp4_pih_outcome <- wp4_pih_outcome %>%
  mutate(
    PIH_OUTCOME = if_else(PIH_OUTCOME_NIMATS == 1 |
                            PIH_OUTCOME_ICD == 1 |
                            PIH_OUTCOME_MED == 1,
                          1, 0
                          )
    )
```

```{r}
sum(wp4_pih_outcome$PIH_OUTCOME_NIMATS)
sum(wp4_pih_outcome$PIH_OUTCOME_ICD, na.rm=T)
sum(wp4_pih_outcome$PIH_OUTCOME_MED, na.rm=T)
sum(wp4_pih_outcome$PIH_OUTCOME, na.rm=T)
sum(wp4_pih_outcome$PIH_OUTCOME_ICD==1 | wp4_pih$PIH_OUTCOME_MED==1, na.rm=T)
```

```{r}
sum(is.na(wp4_pih_outcome$STUDY_ID))
sum(is.na(wp4_pih_outcome$PREG_ID))
```

```{r}
wp4_pih_outcome <- wp4_pih_outcome %>% select(-STUDY_ID)
```

```{r}
#saving wide dataset of binary flags for pih
write_csv(wp4_pih_outcome, file=paste0("S:/HBS_E062/3. DATA/WP4 Polypharmacy/wp4_pih_outcome_", format(lubridate::now(), "%Y.%m.%d_%H.%M"),".csv"))
```

############################################################################
#######                         REMINDERS/ TO DO                     #######
############################################################################

############################################################################



######################################
#######           END          #######
######################################