---
title: "MuM_PreDiCT WP4 Polypharmacy (2) Medication Exposure Windows"
output: html_notebook
---

last updated: 2023.05.10 (Lisa Kent)

This notebook:
* prepares full BNF codes
* for every instance of a dispensed item, creates a pregnancy level flag if medication issued within the exposure window 1)90 days pre-preg 2)trimester 1

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```
__Lode odbc (Open Database Connectivity) package__
```{r}
if(!require(odbc)) install.packages("odbc")
library(odbc)
```

__Declare connection to database__
```{r}
con <- dbConnect(odbc::odbc(), "hirusql3", timeout = 10)
```



__READ IN EPD FROM DATABASE__
```{r}
#likely to be very large!!!
epd <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[EPD_1]")
wp4_epd <- epd
```

```{r}
#clean variable names and class
names(wp4_epd) <- toupper(names(wp4_epd))
wp4_epd$ISSUE_DATE <- as.Date(wp4_epd$ISSUE_DATE)
glimpse(wp4_epd)
```

__Filter date range for study__
```{r}
# keep from Jan 2010 to Dec 2021

wp4_epd <- wp4_epd %>% 
  filter(ISSUE_DATE >= as.Date("2010-01-01") & ISSUE_DATE <= as.Date("2021-12-31"))

```

```{r}
#sanity check - do we have correct dates?
wp4_epd %>% group_by(ISSUE_DATE) %>% count()
wp4_epd %>% filter (ISSUE_DATE >= as.Date("2021-01-01")) %>% group_by(ISSUE_DATE) %>% count()

#are there any missing dates?
sum(is.na(wp4_epd$ISSUE_DATE))
```


```{r}
n_distinct(wp4_epd$VMP_NM)
n_distinct(wp4_epd$VTM_NM)
n_distinct(wp4_epd$BNF_CHAPTER)
n_distinct(wp4_epd$BNF_CHAPTER, wp4_epd$BNF_SECTION)
n_distinct(wp4_epd$BNF_CHAPTER, wp4_epd$BNF_SECTION, wp4_epd$BNF_PARAGRAPH)
n_distinct(wp4_epd$BNF_CHAPTER, wp4_epd$BNF_SECTION, wp4_epd$BNF_PARAGRAPH, wp4_epd$BNF_SUB_PARAGRAPH)
```



__CREATE FLAGS IF MEDICATION ISSUED WITHIN EXPOSURE WINDOWS__
```{r}
epd_exposure <- left_join(wp4_epd, wp4_dates, by = "STUDY_ID", relationship = "many-to-many")
```

```{r}
#flag each item if it was issued during the exposure windows
epd_exposure <- epd_exposure %>%
  mutate(
    MED_EXPOSURE_PREPREG = if_else(ISSUE_DATE >= DATE_START_MINUS_90 & ISSUE_DATE < LMP_CALC, 1, 0),
    MED_EXPOSURE_TRIM1 = if_else(ISSUE_DATE >= LMP_CALC & ISSUE_DATE < T1_END, 1, 0)
  )

#sanity check
epd_exposure %>%
  group_by(MED_EXPOSURE_PREPREG) %>% count()

epd_exposure %>%
  group_by(MED_EXPOSURE_TRIM1) %>% count()
```


```{r}
epd_exposure <- epd_exposure %>% filter (!(is.na(DATE_START_MINUS_90)))
```

```{r}
levels(as.factor(epd_exposure$BNF_CHAPTER))
levels(as.factor(epd_exposure$BNF_SECTION))
levels(as.factor(epd_exposure$BNF_PARAGRAPH))
levels(as.factor(epd_exposure$BNF_SUB_PARAGRAPH))
```

```{r}
#Add in 0 in front of single digit codes to allow for a full 8 digit bnf code to be created
epd_exposure$BNF_CHAPTER <- as.character(epd_exposure$BNF_CHAPTER)
epd_exposure$BNF_SECTION <- as.character(epd_exposure$BNF_SECTION)
epd_exposure$BNF_PARAGRAPH <- as.character(epd_exposure$BNF_PARAGRAPH)
epd_exposure$BNF_SUB_PARAGRAPH <- as.character(epd_exposure$BNF_SUB_PARAGRAPH)


epd_exposure <- epd_exposure %>%
  mutate(
    BNF_CHAPTER = if_else(nchar(BNF_CHAPTER) == 1, paste0("0", BNF_CHAPTER), BNF_CHAPTER),
    BNF_SECTION = if_else(nchar(BNF_SECTION) == 1, paste0("0", BNF_SECTION), BNF_SECTION),
    BNF_PARAGRAPH = if_else(nchar(BNF_PARAGRAPH) == 1, paste0("0", BNF_PARAGRAPH), BNF_PARAGRAPH),
    BNF_SUB_PARAGRAPH = if_else(nchar(BNF_SUB_PARAGRAPH) == 1, paste0("0", BNF_SUB_PARAGRAPH), BNF_SUB_PARAGRAPH),
    BNF_CODE = paste0(BNF_CHAPTER, BNF_SECTION, BNF_PARAGRAPH, BNF_SUB_PARAGRAPH)
         )


levels(as.factor(epd_exposure$BNF_CHAPTER))
levels(as.factor(epd_exposure$BNF_SECTION))
levels(as.factor(epd_exposure$BNF_PARAGRAPH))
levels(as.factor(epd_exposure$BNF_SUB_PARAGRAPH))

```


```{r}
levels(as.factor(epd_exposure$BNF_CODE))
```

```{r}
# COMPARISON OF BNF_CODES and MEDICATION NAMES IN NI DATA VS. CPRD

# read in the BNF code and name for the 570 exposure medications
wp4_bnf_codes <- read_csv("S:/HBS_E062/3. DATA/WP4 Polypharmacy/BNF codes 2023.05.15.csv",
                          col_types = list(col_character()))
# any with a zero at start are being concatenated to 7 chars

wp4_bnf_codes <- wp4_bnf_codes %>%
  mutate(bnf_code = if_else(nchar(as.character(bnf_code)) == 7, paste0("0", bnf_code), bnf_code))
wp4_bnf_codes$WP4_FLAG <- 1

epd_codes1 <- epd_exposure %>% group_by(BNF_CODE) %>%
  summarise(VTM_NM_LIST = paste(sort(unique(VTM_NM)), collapse="; "))

epd_codes2 <- epd_exposure %>% group_by(BNF_CODE) %>% summarise(COUNT = n_distinct(PREG_ID))
epd_codes2$EPD_FLAG <- 1

epd_codes3 <- left_join(epd_codes1, epd_codes2, by = "BNF_CODE")

code_compare <- full_join(epd_codes3, wp4_bnf_codes, by = c("BNF_CODE" = "bnf_code"))

code_compare <- code_compare %>% arrange(BNF_CODE)

code_compare <- code_compare %>% mutate(BNF_CODE = paste0("BNF_", BNF_CODE))

code_compare %>% group_by(EPD_FLAG, WP4_FLAG) %>% count()

write_csv(code_compare, file = paste0("S:/HBS_E062/3. DATA/WP4 Polypharmacy/code_compare_", format(lubridate::now(), "%Y.%m.%d_%H.%M"),".csv"))
# 
# epd_exposure <- left_join(epd_exposure, code_compare, by = "BNF_CODE")
# 
# #inspect bnf codes in epd that have no link to mum-predict bnf codes
# #ensure that they are for non-medicinal items (e.g. dressings)
# 
# epd_exposure %>% filter(is.na(description)) %>% group_by(BNF_CODE, VTM_NM) %>% count()
# epd_exposure %>% filter(is.na(description)) %>% group_by(BNF_CODE) %>% count()
# epd_exposure %>% filter(is.na(description)) %>% group_by(BNF_CODE) %>% summarise(COUNT = n_distinct(PREG_ID))
```


__Create separate "wide" data frame for 90 days pre-preg exposure__
```{r}
#pre-pregnancy exposures
epd_prepreg <- epd_exposure %>% filter(MED_EXPOSURE_PREPREG == 1) %>% select(PREG_ID, BNF_CODE)
```

```{r}
#sanity check - does filtering by 90day pre-preg exposure reduce the number of items?
levels(as.factor(epd_prepreg$BNF_CODE))
```



```{r}
#flip to wide (i.e. each BNF_CODE becomes a binary variable - 1=exposed; 0=not)
epd_prepreg <- epd_prepreg %>%
  group_by(PREG_ID, BNF_CODE) %>%
  summarise(COUNT = n())

epd_prepreg$COUNT[epd_prepreg$COUNT>0] <- 1

epd_prepreg_wide <- epd_prepreg %>%
  pivot_wider(names_from = BNF_CODE,
              values_from = COUNT,
              values_fill = 0)

head(epd_prepreg_wide)
```


__Create separate "wide" data frame for trimester 1 exposure__
```{r}
epd_trim1 <- epd_exposure %>% filter(MED_EXPOSURE_TRIM1 == 1) %>% select(PREG_ID, BNF_CODE)
```

```{r}
#sanity check - does filtering by trimester 1 exposure reduce the number of items?
levels(as.factor(epd_trim1$BNF_CODE))
```



```{r}
#flip to wide (i.e. each BNF_CODE becomes a binary variable - 1=exposed; 0=not)
epd_trim1 <- epd_trim1 %>%
  group_by(PREG_ID, BNF_CODE) %>%
  summarise(COUNT = n())

epd_trim1$COUNT[epd_trim1$COUNT>0] <- 1

epd_trim1_wide <- epd_trim1 %>%
  pivot_wider(names_from = BNF_CODE,
              values_from = COUNT,
              values_fill = 0)

head(epd_trim1_wide)
```

```{r}
#add prefix to BNF_CODE to allow for easier variable selection later in pipeline
epd_prepreg_wide <- epd_prepreg_wide %>% rename_with(~ paste0("BNF_", .), -PREG_ID)

epd_trim1_wide <- epd_trim1_wide %>% rename_with(~ paste0("BNF_", .), -PREG_ID)
```



```{r}
#saving wide dataset of binary flags for exposure to medications (Full BNF code) in 90 days pre-pregnancy
write_csv(epd_prepreg_wide, file=paste0("S:/HBS_E062/3. DATA/WP4 Polypharmacy/epd_prepreg_wide_", format(lubridate::now(), "%Y.%m.%d_%H.%M"),".csv"))

#saving wide dataset of binary flags for exposure to medications (Full BNF code) in trimester one
write_csv(epd_trim1_wide, file=paste0("S:/HBS_E062/3. DATA/WP4 Polypharmacy/epd_trim1_wide_", format(lubridate::now(), "%Y.%m.%d_%H.%M"),".csv"))

```


############################################################################
#######                         REMINDERS/ TO DO                     #######
############################################################################

############################################################################



######################################
#######           END          #######
######################################